---
title: "history-match-waves"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{history-match-waves}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(imptools)
library(lhs)
library(DiceKriging)
library(emtools)
library(viztools)

```

A set of toy model Functions
```{r toy}

# Friedman
friedman = function(x){
  
  stopifnot(length(x)==4)
  
  out = 10*sin(pi*x[1]*x[2]) + 20*(x[3] - 0.5)^2 + 10*x[4] # + 5*x5
  out
}

park1 = function(x){
  
  stopifnot(length(x)==4)
  
  out = x[1]/2 * sqrt( 1+ (x[2]+ (x[3]^2) * (x[4] / x[1]^2))) + x[1] + 3*x[4]*exp(1+sin(x[3]))
  out
}

park2 = function(x){
  
  stopifnot(length(x)==4)
  
  out = (2/3 * exp(x[1]+x[2])) - (x[4] *sin(x[3])) + x[3]
  out
}


run_model = function(X){
  # Generates multivariate output from a design matrix, using
  # the friedman, park1 and park2 functions.
  # Input is a design matrix with 4 columns on the [0,1] cube
  # Output is a 3 column matrix of model output
  stopifnot(ncol(X)==4)
  n = nrow(X)
  
  Y = matrix(NA, nrow = n, ncol = 3)
  colnames(Y) = c('y1', 'y2', 'y3')
  
  for(i in 1:n){
    
    x = X[i, ]
    
    y1 = friedman(x)
    y2 = park1(x)
    y3 = park2(x)
    
    Y[i, ] = c(y1, y2, y3)
    
  }
  Y
}
```

### initial design  
Set up an initial design, sampling sparsely from the input domain (as if the model was expensive to run).  
Experiment with n (initial design points), n.app (number of points to augment the design with in each wave), and uncertainty budget.
```{r}
n = 10 # number of initial design points
k = 4  # number of model inputs
d = 3  # number of model outputs

# simplest method is to choose a subset of the discovered NROY
# points to append to the design
  
n_app = 10 # number of points to append

  X = maximinLHS(n, k = k,dup = 2)
colnames(X) <- c('x1', 'x2', 'x3', 'x4')

#X_target = matrix(runif(4), nrow = 1) # uncomment this and comment the next line for a random target input.
X_target = matrix(rep(0.25,4), nrow = 1)

# run the model outside of the function
#Y = run_model(X)
#Y_target = run.model(X_target)

```

```{r}

Y_raw = run_model(X)
Y_target_raw = run_model(X_target)

Y_all_norm = normalize(rbind(Y_raw, Y_target_raw))
Y = Y_all_norm[1:n, ]
Y_target = matrix(Y_all_norm[n+1, ], nrow = 1)
```


```{r}
# Assume the entire uncertainty budget is in the observational uncertainty initially
#obs.sd.list = list(0.013,0.013,0.013)
# at around 0.01, the HM breaks, putting the I of the target point above 3  
obs_sd_list = list(0.015,0.015,0.015)
disc_list = list(0,0,0) 
disc_sd_list = list(0, 0, 0) 

thres = 3   # implausibility threshold

mins_aug = rep(0,4)
maxes_aug = rep(1,4)

# Improvements to the above function from Andrianakis et al. (2015)
# 1) Reduce the range the emulator is fit over, as the waves continue.
# 2) Sample candidate design points from NEAR the existing NROY design points.

```

This function augments the original design with points that are found to be NROY and well spaced out using the emulator.
It samples n_aug points uniformly across the original space, a large number of times, and finds those that are NROY. It then chooses n_app of these and finds the minumum distance. It repeats this nreps times, and finds the set of n_app points with the largest minumum distance.

```{r}
wave1 = addNroyDesignPoints(X = X,
                            Y = Y,
                            Y_target = Y_target,
                            n_aug = 100000,
                            nreps = 1000,
                            n_app = n_app,
                            mins_aug = mins_aug,
                            maxes_aug = maxes_aug,
                            thres = 3,
                            disc_list = disc_list,
                            disc_sd_list = disc_sd_list,
                            obs_sd_list = obs_sd_list)
```

```{r, fig.width = 8, fig.height = 8}
 pairs(rbind(wave1$X_nroy, wave1$X, wave1$X_out, wave1$X_mm, X_target),
       col = c(rep('darkgrey', nrow(wave1$X_nroy)),
               rep('black', nrow(wave1$X)),
               rep('blue', nrow(wave1$X_out)),
               rep('red', nrow(wave1$X_mm)),
               'gold'
               ),
       pch = 19,
       xlim = c(0,1),
       ylim = c(0,1)
      )
```

This part takes the n.aug points found to be NROY, and chooses n_app of them that are found to be furthest apart. These are appended to the design points that are inside the marginal NROY ranges.
```{r, fig.width = 8, fig.height = 8}

# This part takes the n.aug points found to be NROY, and chooses n.app of them that are found to be
# furthest apart. These are appended to the design points that are inside the marginal NROY ranges.
#X2 = rbind(X[keep1_ix, ] , ChooseMaximinNroy(n_app = n_app, waveobj=wave1, nreps = 10000))

# Run the model at the new design
Y2_raw = run_model(wave1$X_out)
Y2 = normalize(Y2_raw, wrt = Y_raw)

wave2 = addNroyDesignPoints(X = wave1$X_out,
                               Y = Y2,
                               Y_target = Y_target,
                               n_aug = 100000,
                               nreps = 1000,
                              n_app = n_app,
                               mins_aug = mins_aug,
                               maxes_aug = maxes_aug,
                               thres = 3,
                               disc_list=disc_list,
                               disc_sd_list = disc_sd_list,
                               obs_sd_list = obs_sd_list)
```

```{r, fig.width = 8, fig.height = 8}

 pairs(rbind(wave2$X_nroy, wave2$X, wave2$X_out, wave2$X_mm, X_target),
       col = c(rep('darkgrey', nrow(wave2$X_nroy)),
               rep('black', nrow(wave2$X)),
               rep('blue', nrow(wave2$X_out)),
               rep('red', nrow(wave2$X_mm)),
               'gold'
               ),
       pch = 19,
       xlim = c(0,1),
       ylim = c(0,1)
      )
```



```{r}
Y3_raw = run_model(wave2$X_out)
Y3 = normalize(Y3_raw, wrt = Y_raw)

wave3 = addNroyDesignPoints(X = wave2$X_out,
                               Y = Y3,
                               Y_target = Y_target,
                               n_aug = 100000,
                               nreps = 1000,
                              n_app = n_app,
                               mins_aug = mins_aug,
                               maxes_aug = maxes_aug,
                               thres = 3,
                               disc_list=disc_list,
                               disc_sd_list = disc_sd_list,
                               obs_sd_list = obs_sd_list)
```

```{r, fig.width = 8, fig.height = 8}

 pairs(rbind(wave3$X_nroy, wave3$X, wave3$X_out, wave3$X_mm, X_target),
       col = c(rep('darkgrey', nrow(wave3$X_nroy)),
               rep('black', nrow(wave3$X)),
               rep('blue', nrow(wave3$X_out)),
               rep('red', nrow(wave3$X_mm)),
               'gold'
               ),
       pch = 19,
       xlim = c(0,1),
       ylim = c(0,1)
      )
```


```{r loop}
loopAddDesignPoint = function(X, 
                              Y, 
                              Y_target, 
                              n_aug,
                              mins_aug,
                              maxes_aug,
                              thres = 3, 
                              disc_list,
                              disc_sd_list,
                              obs_sd_list,
                              waves = 3,
                              n_app = n_app,
                              nreps = 1000,
                              shrink = TRUE){
  
  if(length(n_aug) ==1){
    n_aug = rep(n_aug, waves)
  }
  
  wavelist = vector(mode='list', length = waves)
  Xlist  = vector(mode='list', length = waves)
  Ylist  = vector(mode='list', length = waves)
  
  for(i in 1:waves){
    
    wave = addNroyDesignPoints(X = X, 
                                  Y = Y, 
                                  Y_target = Y_target,
                                  n_aug = n_aug[i], 
                                  mins_aug = mins_aug,
                                  maxes_aug = maxes_aug,
                                  thres = thres,
                               nreps = nreps,
                                  disc_list=disc_list,
                                  disc_sd_list = disc_sd_list,
                                  obs_sd_list = obs_sd_list)
    
    # This part excludes any original design points outside of the ranges returned as NROY by
    # add_nroy_design_points. This shrinks down the domain that the emulator
    # is fit over.
    
    if(shrink){
      keep_ix = which(apply(X, FUN = withinRange,1,
                             maxes = wave$X_nroy_max,
                             mins = wave$X_nroy_min))
      
      # This part takes the n_aug points found to be NROY, 
      # and chooses n_app of them that are found to be furthest apart. 
      # These are appended to the design points that are inside the marginal NROY ranges. 
      
      Xnew = rbind(X[keep_ix, ] , chooseMaximinNroy(n_app = n_app, waveobj=wave, nreps = nreps))
      
    }
    
    else{
      Xnew = rbind(X, chooseMaximinNroy(n_app = n_app, waveobj = wave, nreps = nreps) )
    }
    
    # Run the model at the new design
    Ynew_raw = run_model(Xnew)
    
    # Normalise with respect to the original output.
    Ynew = normalize(Ynew_raw, wrt = Ynew_raw)
    
    X = Xnew
    Y = Ynew
    
    wavelist[[i]] = wave
    Xlist[[i]] = X
    Ylist[[i]] = Y
    
  }
  
  return(list(wavelist = wavelist, shrink = shrink, Xlist = Xlist, Ylist = Ylist, nreps = nreps))
}

```


This loops through model runs, shrinking the NROY space each time
```{r, echo = FALSE, results = 'hide', message=FALSE, hide=TRUE}
# # this is hanging
# waves = 3
# n_aug = rep(100000, waves)
# # Increase n_aug as you go through the waves?
# #n_aug = c(100000, 100000, 100000, 100000, 100000, 100000, 100000, 100000)
# 
# wavetest = loopAddDesignPoint (X = X, 
#                                Y = Y,
#                                Y_target = Y_target,
#                                n_aug = n_aug, 
#                                nreps = 1000,
#                                mins_aug = mins_aug,
#                                maxes_aug = maxes_aug,
#                                thres = 3,
#                                disc_list=disc_list,
#                                disc_sd_list = disc_sd_list,
#                                obs_sd_list = obs_sd_list,
#                            n_app = n_app,
#                            waves = waves,
#                            shrink = TRUE
# )

```


```{r, width = 8, height = 8}

makeTransparent = function(someColor, alpha=100)
  # Transparent colours for plotting
{
  newColor<-col2rgb(someColor)
  apply(newColor, 2, function(curcoldata){rgb(red=curcoldata[1], green=curcoldata[2],
                                              blue=curcoldata[3],alpha=alpha, maxColorValue=255)})
}

reset <- function() {
  # Allows annotation of graphs, resets axes
  par(mfrow=c(1, 1), oma=rep(0, 4), mar=rep(0, 4), new=TRUE)
  plot(0:1, 0:1, type="n", xlab="", ylab="", axes=FALSE)
}


# for(i in 1:length(test$wavelist)){
#   
#   if(i==1){Xem = X}
#   else{Xem = test$Xlist[[i-1]]}
#   
#   x = rbind(test$wavelist[[i]]$X_nroy, Xem, test$Xlist[[i]], X_target)
#   
#    # png(file = paste0('graphics/anim/toy_hm_waves_01_',sprintf("%04d",i),'.png'),
# #width = 480, height = 480)
#            
#   par(oma = c(0,0,0,3), bg = 'white')
#   cols = c(rep('darkgrey', nrow(test$wavelist[[i]]$X_nroy)),
#            rep('blue', nrow(Xem)),
#            rep('gold', nrow(test$Xlist[[i]])),
#           'red')
#  
#   pairs(x,
#         gap = 0, lower.panel = NULL, xlim = c(0,1), ylim = c(0,1),
#         cex.labels = 3,
#         col.axis = 'white',
#         col = cols,
#         pch = 19,
#         cex = 1
#            )
#   
#   reset()
# legend('bottom',
#        legend = c('NROY samples', 'Design points', 'test', 'True input (hidden)'),
#        col = c('darkgrey', 'blue', 'gold', 'red'),
#        inset = 0.05,
#        cex = 1.1,
#        pch = 19
#        )
# #dev.off()
# # use 
# # convert -delay 100 toy*.png -loop 0 movie.gif
# # to animate the pngs.
# }

```
